\pdftitle

### 2017-05-17 maxNt.tex
new material on projection

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-17T15:27:36+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

We show two examples of this discrepancy in
\figs~\ref{fig:diff_maxent_pop_sample}
and~\ref{fig:diff_maxent_pop_sample_realdata}. In the former the
constraints are $\expe{\yxs} = c_1$ and
$\expe{\yxxs} = c_2$, with

Suppose our desired constraints are $\expe{\yxs } = c_1$ and
$\expe{\yxxs} = c_2$, the expectations being given by the fourth term of
\eqn~\eqref{eq:expe_products}. Applying \me\ to the distribution of the
average of the full population we find
\begin{equation}
  \label{eq:full_pop_av_example_ME}
  \p(\yXf = \yRf \| \yHa)  = \yL\,
\binom{N}{N \yRf}\exp\biggl[
\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr]
\end{equation}
with such parameters $\yL$, $\yL_1$, $\yL_2$ as to satisfy normalization
and constraints. We call this state of knowledge $\yHa$. Marginalization by
\eqn~\eqref{eq:marginal} then gives
\begin{equation}
  \label{eq:full_pop_av_example_ME_marg}
  \p(\yxs = \yrs \| \yHa)  = \yL
  \sum_{N\yRf=0}^N
  \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr].
\end{equation}
Applying \me\ to the distribution of the average of the sample, using the
third term of \eqn~\eqref{eq:expe_products}, we find
\begin{equation}
  \label{eq:sample_av_example_ME}
  \p(\yxs = \yrs \| \yHb)  = \yl\,
  \binom{n}{n\yrs}
  \exp\biggl[\yl_2 \yrs \,\frac{n \yrs-1}{n-1} + \yl_1\, \yrs\biggr]
\end{equation}
with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
We call this state of knowledge $\yHb$.
%\setlength{\intextsep}{0.5ex}%
%\begin{wrapfigure}{r}{0.5\linewidth}
\begin{figure}[!t]
\centering
\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
\caption{Plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
  level followed by sample marginalization (blue triangles),
  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
\label{fig:diff_maxent_pop_sample}
\end{figure}%maxent_pop_or_sample.nb
% \end{wrapfigure}
\begin{figure}[!t]
\centering
\includegraphics[width=0.99\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
% 
\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
  the population level followed by sample marginalization (blue triangles),
  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
  $\expe{\sav{\yx\yx\yx}}=1.48\times10^{-4}$,
  $\expe{\sav{\yx\yx\yx\yx}}=8.81\times 10^{-6}$.}
\label{fig:diff_maxent_pop_sample_realdata}
\end{figure}%maxent_pop_or_sample.nb

@@

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-17T15:41:05+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

%\setlength{\intextsep}{0.5ex}%
%\begin{wrapfigure}{r}{0.5\linewidth}
\begin{figure}[!t]
\centering
\includegraphics[width=0.75\linewidth]{different_maxent_pop_sample_200_realdata_2mom.pdf}%
\caption{Plots of $\p(\yxs = \yrs)$ constructed by \me\ at the population
  level followed by sample marginalization (blue triangles),
  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$.}
\label{fig:diff_maxent_pop_sample}
\end{figure}%maxent_pop_or_sample.nb
% \end{wrapfigure}
\begin{figure}[!t]
\centering
\includegraphics[width=0.75\linewidth]{different_maxent_pop_sample_200_realdata_4mom.pdf}
% 
\caption{Linear and log-plots of $\p(\yxs = \yrs)$ constructed by \me\ at
  the population level followed by sample marginalization (blue triangles),
  \eqn~\eqref{eq:full_pop_av_example_ME_marg}, and at the sample level (red
  circles), \eqn~\eqref{eq:sample_av_example_ME}. In this case $N=5000$,
  $n=200$, $\expe{\yxs}=0.0478$, $\expe{\yxxs}=0.00257$,
  $\expe{\sav{\yx\yx\yx}}=1.48\times10^{-4}$,
  $\expe{\sav{\yx\yx\yx\yx}}=8.81\times 10^{-6}$.}
\label{fig:diff_maxent_pop_sample_realdata}
\end{figure}%maxent_pop_or_sample.nb
We show two examples of this discrepancy in
\figs~\ref{fig:diff_maxent_pop_sample}
and~\ref{fig:diff_maxent_pop_sample_realdata}, for $N=5000$, $n=200$. In
the first example the constraints are $\expe{\yxs} = c_1$ and
$\expe{\yxxs} = c_2$, with $c_1=0.0478$ and $c_2=0.00257$. The distribution
obtained at the sample level is broader than the one obtained at the
population level; the tails of the two distributions are very different.
The second example includes two additional constraints
$\expe{\sav{\yx\yx\yx}}= c_3$, $\expe{\sav{\yx\yx\yx\yx}}=c_4$ with
$c_3=0.000148$, $c_4=8.81\times 10^{-6}$. The distribution obtained at the
population level has two modes, replaced by only one in the distribution
obtained at the sample level; the tails are very different also in this
case. The constraints used in these examples have neurobiologically
realistic values \cite{rostamietal2016}.

@@

  *******

Suppose our desired constraints are $\expe{\yxs } = c_1$ and
$\expe{\yxxs} = c_2$, the expectations being given by the fourth term of
\eqn~\eqref{eq:expe_products}. Applying \me\ to the distribution of the
average of the full population we find
\begin{equation}
  \label{eq:full_pop_av_example_ME}
  \p(\yXf = \yRf \| \yHa)  = \yL\,
\binom{N}{N \yRf}\exp\biggl[
\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr]
\end{equation}
with such parameters $\yL$, $\yL_1$, $\yL_2$ as to satisfy normalization
and constraints. We call this state of knowledge $\yHa$. Marginalization by
\eqn~\eqref{eq:marginal} then gives
\begin{equation}
  \label{eq:full_pop_av_example_ME_marg}
  \p(\yxs = \yrs \| \yHa)  = \yL
  \sum_{N\yRf=0}^N
  \binom{n}{n\yrs} \binom{N-n}{N\yRf-n\yrs} \exp\biggl[
\yL_2 \yRf \,\frac{N \yRf-1}{N-1} + \yL_1\, \yRf\biggr].
\end{equation}
Applying \me\ to the distribution of the average of the sample, using the
third term of \eqn~\eqref{eq:expe_products}, we find
\begin{equation}
  \label{eq:sample_av_example_ME}
  \p(\yxs = \yrs \| \yHb)  = \yl\,
  \binom{n}{n\yrs}
  \exp\biggl[\yl_2 \yrs \,\frac{n \yrs-1}{n-1} + \yl_1\, \yrs\biggr]
\end{equation}
with such parameters $\yl$, $\yl_1$, $\yl_2$ as to satisfy the constraints.
We call this state of knowledge $\yHb$.
